Containers and R---the Rockerverse and beyond

### 2020-01-13 manuscript.Rmd
update headlines (no more assignments to people) and improve package intro

## Interfaces for Docker in R

Interfacing with the Docker daemon is typically done through the Docker Command Line Interface (CLI).
However, moving back and forth between an R console and the command line can create friction in workflows and reduce reproducibility.
A number of first-order R packages provide a interface to the Docker CLI, allowing to automate interaction with the Docker CLI from an R console.

Each of these packages has particular advantages as they provide function wrappers for interacting with the Docker CLI at different stages of a container's life cycle. Examples of such interactions are installing the Docker software, creating `Dockerfile`s, building an image, and launching a container on a local machine or on the cloud. As such, the choice of which package is most useful depends on the use-case at hand.
<!-- ^ in the examples listed above, the name of the packages which provide such functionality could be mentioned, this would highlight the variety in functionality already. -->

@@

**stevedore** ([https://cran.r-project.org/package=stevedore](https://cran.r-project.org/package=stevedore)) 
<!-- from the CRAN description:
Work with containers over the Docker API. Rather than using system calls to interact with a docker client, using the API directly means that we can receive richer information from docker. The interface in the package is automatically generated using the 'OpenAPI' (a.k.a., 'swagger') specification, and all return values are checked in order to make them type stable.
-->

**dockyard** ([https://github.com/thebioengineer/dockyard](https://github.com/thebioengineer/dockyard)) has the goal of lowering barrier to creating `Dockerfile`s and Docker images, and deploying Docker containers. 
The way docker images are made is by reading and following a set of instructions found in `Dockerfile`s.
This requires stepping out of the R ecosystem and into an infrastructure mindset where the user needs to specify the environment they require.
Because `Dockerfile`s can be read as a series of steps, this lends itself well to the `tidyverse` style of programming where each step is piped into the next.
dockyard has a `Dockerfile` object that either starts out as an empty `Dockerfile`, or can use existing `Dockerfile` to use as a template.
This combined with a series of functions that are analogous to the steps commonly found in a `Dockerfile`, it makes building Docker images much more approachable to R users.
There are also wrappers for commonly used steps for R programmers like adding libraries and copying files into the R Docker containers.
Once the `Dockerfile` is created, dockyard has the functionality to build the image and deploy it.

@@

- [`dockerfiler`](https://github.com/ColinFay/dockerfiler/)
- [`containerit`](https://github.com/o2r-project/containerit/)
- [`dockr`](https://github.com/smaakage85/dockr), see also http://smaakage85.netlify.com/2019/12/21/dockr-easy-containerization-for-r/
- [`dockertest`](https://github.com/traitecoevo/dockertest/)
- [`renv` with Docker](https://rstudio.github.io/renv/articles/docker.html)

@@

## Deployment to cloud services [div]

@@

## Using R to power enterprise software in production environments [@jnolis, @nolistic]

@@

## Deployment to the cloud [@hongooi]

@@

## Using Docker in a Data Science Platform

>>>>

## Interfaces for Docker in R
\label{interfaces}

Interfacing with the Docker daemon is typically done through the [Docker Command Line Interface](https://docs.docker.com/engine/reference/commandline/cli/) (Docker CLI).
However, moving back and forth between an R console and the command line can create friction in workflows and reduce reproducibility.
A number of first-order R packages provide a interface to the Docker CLI, allowing to automate interaction with the Docker CLI from an R console.

Each of these packages has particular advantages as they provide function wrappers for interacting with the Docker CLI at different stages of a container's life cycle. Examples of such interactions are installing the Docker software, creating `Dockerfile`s (\pkg{dockerfiler}), \pkg{containerit}), building images and launching a containers (\pkg{stevedore}, \pkg{docker}) on a local machine or on the cloud.
As such, the choice of which package is most useful depends on the use-case at hand, but also the users level of expertise.

@@

**\pkg{stevedore}** ([https://cran.r-project.org/package=stevedore](https://cran.r-project.org/package=stevedore)) 
<!-- from the CRAN description:
Work with containers over the Docker API. Rather than using system calls to interact with a docker client, using the API directly means that we can receive richer information from docker. The interface in the package is automatically generated using the 'OpenAPI' (a.k.a., 'swagger') specification, and all return values are checked in order to make them type stable.
-->

**\pkg{AzureContainers}** ([https://cran.r-project.org/package=AzureContainers](https://cran.r-project.org/package=AzureContainers)) is an interface to a number of container-related services in Microsoft's [Azure Cloud](https://azure.microsoft.com/), namely [Container Instances](https://azure.microsoft.com/en-us/services/container-instances/), [Container Registry](https://azure.microsoft.com/en-us/services/container-registry/), and [Kubernetes Service](https://azure.microsoft.com/en-us/services/kubernetes-service/).
While it is mainly intended for working with Azure, as a convenience feature it includes lightweight, cross-platform shells to Docker and Kubernetes (tools `kubectl` and `helm`).
These can be used to create and manage arbitrary Docker images and containers, as well as Kubernetes clusters on any platform or cloud service.

**\pkg{babelwhale}** ([https://cran.r-project.org/package=babelwhale](https://cran.r-project.org/package=babelwhale)) 
allows executing and interacting with containers, which can use either Docker or Singularity as a backend.
The package provides a unified interface to interact with Docker and Singularity containers.
Users can, for example, execute a command inside a container, mount a volume or copy a file.

**\pkg{harbor}** ([https://github.com/wch/harbor](https://github.com/wch/harbor)
<!-- from the github description
is an R package for controlling docker containers on local and remote hosts.
-->

**\pkg{dockyard}** ([https://github.com/thebioengineer/dockyard](https://github.com/thebioengineer/dockyard)) has the goal of lowering barrier to creating `Dockerfile`s and Docker images, and deploying Docker containers. 
The way docker images are made is by reading and following a set of instructions found in `Dockerfile`s.
This requires stepping out of the R ecosystem and into an infrastructure mindset where the user needs to specify the environment they require.
Because `Dockerfile`s can be read as a series of steps, this lends itself well to the `tidyverse` style of programming where each step is piped into the next.
dockyard has a `Dockerfile` object that either starts out as an empty `Dockerfile`, or can use existing `Dockerfile` to use as a template.
This combined with a series of functions that are analogous to the steps commonly found in a `Dockerfile`, it makes building Docker images much more approachable to R users.
There are also wrappers for commonly used steps for R programmers like adding libraries and copying files into the R Docker containers.
Once the `Dockerfile` is created, dockyard has built-in functionality to build the image and run a container.

@@

- [`dockerfiler`](https://github.com/ColinFay/dockerfiler/)
- [`containerit`](https://github.com/o2r-project/containerit/)
- [`dockr`](https://github.com/smaakage85/dockr), see also http://smaakage85.netlify.com/2019/12/21/dockr-easy-containerization-for-r/
- [`dockertest`](https://github.com/traitecoevo/dockertest/)
- [`renv` with Docker](https://rstudio.github.io/renv/articles/docker.html)
- https://github.com/gigantum/base-images

@@

## Deployment to cloud services

@@

## Using R to power enterprise software in production environments

@@

## Deployment to the cloud

@@

## Using Docker in a commercial Data Science Platform
